---
title: "BIS"
output:
  github_document:
    fig_width: 8
    fig_height: 4
---

The `BIS` package package provides an `R` interface to data hosted by the [Bank for International Settlements](https://www.bis.org), specifically the [single-file data sets](https://www.bis.org/statistics/full_data_sets.htm) available on the BIS homepage.

## Install package

The package can be installed from GitHub.

```{r install, eval=FALSE, message=FALSE, warning=FALSE}
library(devtools)
install_github("stefanangrick/BIS")  # GitHub
```

## Example usage

To import data, first load the package:

```{r loading, message=FALSE, warning=FALSE}
library("BIS")
```

Next, run the `get_datasets()` function to obtain a list of available data sets:

```{r datasets, message=FALSE, warning=FALSE}
datasets <- get_datasets()
head(datasets, 20)
```

The `get_datasets()` function returns a [tibble](https://tibble.tidyverse.org/) data frame listing the available data sets. The column `url` can be used as input for the `get_bis()` function which downloads, parses and imports the corresponding data set.

To import monthly-frequency data on [central banks' policy rates](https://www.bis.org/statistics/cbpol.htm), run:

```{r rates, message=FALSE, warning=FALSE}
rates <- get_bis(datasets$url[datasets$id == "full_cbpol_m_csv"])
head(rates)
```

Parsing may fail if the amount of available memory is insufficient for executing the necessary pivot operation. This is the case with large data sets in particular (e.g. the Locational banking statistics and Debt securities statistics). As a workaround, users may wish to set `auto_pivot = FALSE` when calling `get_bis()`, then subset the data and run `pivot_longer_bis()` manually. See `??get_bis` for an example.

To plot the data using [ggplot2](https://ggplot2.tidyverse.org), run the following:

```{r plot, message=FALSE, warning=FALSE}
library("dplyr")
library("ggplot2")
library("zoo")

rates_plot <- subset(rates, ref_area %in% c("US", "XM", "JP", "GB", "CH", "CA"))
rates_plot <- mutate(rates_plot, date = as.Date(as.yearmon(date, format = "%Y-%m")))

ggplot(rates_plot, aes(date, obs_value, color = reference_area)) +
  geom_line(show.legend = FALSE) +
  facet_wrap(~reference_area) +
  labs(title = "Central bank policy rates",
       subtitle = "% per annum", x = NULL, y = NULL)
```

Note that BIS data sets come with a number of different time formats. The [zoo](https://cran.r-project.org/package=zoo) package (e.g. `as.yearmon()`) should be able to parse most formats.

## Disclaimer

This package is in no way officially related to or endorsed by the [Bank for International Settlements](https://www.bis.org/). It's based on a fork of [CC0](https://cran.r-project.org/src/contrib/Archive/BIS/)-licensed [code by expersso](https://github.com/expersso/BIS). Please don't abuse the BIS's servers with unnecessary calls.
